---
title: "Titanic"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
Author: Gisela Morrone

Set-up of enviroment

```{r}

rm(list = ls())

rmse <- function(pred, obs){  
errores <- pred - obs 
return(sqrt(mean(errores^2)))
}

accuracy <- function(real,predicho){
  mean(real==predicho)
}

##FUNCION PARA RIDGE Y LASSO
funcRIDGE_LASSO <- function(train,testing,parametro, v_exp,v_obj) {

EXPLICATIVAS <- subset(train, select = v_exp)

x <- as.matrix(EXPLICATIVAS[,]) # Removes class

CLASE <- subset(train, select = v_obj)

y <- as.double(as.matrix(CLASE[,])) 

set.seed(3235)

cv.ridge <- cv.glmnet(x, y, family='gaussian', alpha=parametro, standardize=TRUE)

#TEST

EXPLICATIVAS_2 <- subset(testing, select = v_exp)

x_2 <- as.matrix(EXPLICATIVAS_2[,]) # Removes class

ridge_2  <- predict(cv.ridge, newx =x_2, s = "lambda.min") 

result  <-cbind(testing$id,ridge_2)


#Elimino variables auxiliares
rm(EXPLICATIVAS)
rm(x)
rm(CLASE)
rm(y)
rm(cv.ridge)
rm(EXPLICATIVAS_2)
rm(x_2)
rm(ridge_2)
return(result)
}

##ARchive
generarArchivo <-function(dataset,archivo){
#Genera el archivo necesario para subir
#dataset tiene que ser un data.frame que tenga en la primera columna el ID del auto y en la 2da la probabilidad de clase=1
  write.table(data.frame(PassengerId=dataset[,1],
                         Survived=dataset[,2]),
              file=archivo,
              sep=',',
              row.names = F,
              col.names = T
              )
  
}

library(varhandle)
library(dplyr)
library(rpart)
library(rpart.plot)
library(Matrix)
library(foreach)
library(glmnet)

```

Load the data
```{r}
Train <- read.csv(file = "~/Desktop/Data Science/Titanic/train.csv", sep = ",", na.strings = "NULL" ,header = T, stringsAsFactors = F)

Test <- read.csv(file = "~/Desktop/Data Science/Titanic/test.csv", sep = ",",na.strings = "NULL" , header = T, stringsAsFactors = F)


```

Division of Dataset Train in Train and Validation for model, manteining the original proportion of the dataset

```{r}
set.seed(4321)
trProp <- 0.7
trIndex <- c(
             #Cases where objective varible = 1
             sample(which(Train$Survived==1),floor(trProp*sum(Train$Survived==1))),
             #ACases where objective varible = 0
             sample(which(Train$Survived==0),floor(trProp*sum(Train$Survived==0)))
            )
Trn <- Train[trIndex,]
Val <- Train[-trIndex,]
#clean-up
rm(trProp)
```

##Analisis of Dataset Training (Trn)
```{r}
summary(Trn)
as.factor(Trn$Survived)

head(Trn)
```
Analisis of variable SEX > New numeric varible sex, 1 for male, 0 for female
```{r}
table(Trn$Sex)

a <- c(1,"male")
b <- c(2,"female")


sex_cod <- rbind(a,b)
colnames(sex_cod) <- c('sex_cod','Sex')
as.numeric(sex_cod)

Trn <- merge(Trn, sex_cod,by='Sex')

Trn$sex_cod <- unfactor(Trn$sex_cod)

#Check of results
table(Trn$Sex)
table(Trn$sex_cod)

#Elimination of variable Sex
Trn$Sex <-NULL

#clean-up
rm(a)
rm(b)
```

Analisis of variable AGE > Mean age, taking into account Fare, Embarked port, SibSP, Parch, Sex and Class, to imput missing NA

Rest of Missing values with mean according to Class only

```{r}
summary(Trn$Age)

Trn_NNA <- Trn[-is.na(Trn$Age),c("Age","SibSp","Parch","sex_cod","Pclass")]

mean_age_varios <- aggregate(Age ~ ., data =Trn_NNA, FUN=mean)

Trn_2 <- merge(Trn, mean_age_varios,by=c( "SibSp","Parch","sex_cod","Pclass"),all.x = TRUE)

Trn_2$Age.x <- ifelse(is.na(Trn_2$Age.x),Trn_2$Age.y,Trn_2$Age.x )

Trn_2$Age.y <- NULL

colnames(Trn_2 ) <- c("SibSp","Parch","sex_cod","Pclass" ,"PassengerId","Survived" ,"Name" , "Age" ,"Ticket", "Fare","Cabin" ,"Embarked" )

Trn <- Trn_2

#Se reordena el dataset
Trn_2 <- cbind.data.frame(Trn$PassengerId,Trn$Survived,Trn$Pclass,Trn$Name,Trn$sex_cod,Trn$Age,Trn$SibSp,Trn$Parch,Trn$Ticket,Trn$Fare,Trn$Cabin,Trn$Embarked)

colnames(Trn_2 ) <- c("PassengerId","Survived","Pclass"  ,"Name","sex_cod" , "Age" ,"SibSp","Parch","Ticket", "Fare","Cabin" ,"Embarked" )
Trn <- Trn_2

###REVIEW OF RESULTS

summary(Trn$Age)

#Mean according to class
Trn_NNA <- Trn[-is.na(Trn$Age),c("Age","Pclass")]

mean_age_class <- aggregate(Age ~ Pclass, data =Trn, FUN=mean)

Trn_2 <- merge(Trn, mean_age_class,by='Pclass',all.x = TRUE)

Trn_2$Age.x <- ifelse(is.na(Trn_2$Age.x),Trn_2$Age.y,Trn_2$Age.x )

Trn_2$Age.y <- NULL

Trn <- Trn_2

#Se reordena el dataset
Trn_2 <- cbind.data.frame(Trn$PassengerId,Trn$Survived,Trn$Pclass,Trn$Name,Trn$sex_cod,Trn$Age,Trn$SibSp,Trn$Parch,Trn$Ticket,Trn$Fare,Trn$Cabin,Trn$Embarked)

colnames(Trn_2 ) <- c("PassengerId","Survived","Pclass"  ,"Name","sex_cod" , "Age" ,"SibSp","Parch","Ticket", "Fare","Cabin" ,"Embarked" )
Trn <- Trn_2

#clean-up
rm(Trn_2)
rm(Trn_NNA)
```

Analisis of varibale NAME, CABIN, TICKET

```{r}
Trn$Name <-NULL
Trn$Cabin <-NULL
Trn$Ticket <- NULL
```

Analisis of variable Embarked
```{r}
table(Trn$Embarked)

Trn[Trn$Embarked == '',]

#Se asigna el puerto de maxima ocurrencia
Trn$Embarked <- ifelse(Trn$Embarked == '',4,Trn$Embarked)
```

##DATASET VALIDATION
Same convertions done to Train dataset, to Val dataset, without changing parameters learnt
Analisis of variable SEX > New numeric varible sex, 1 for male, 0 for female
```{r}
table(Val$Sex)

Val <- merge(Val, sex_cod,by='Sex')

Val$sex_cod <- unfactor(Val$sex_cod)

#Check of results
table(Val$Sex)
table(Val$sex_cod)

#Elimination of variable Sex
Val$Sex <-NULL

```

Analisis of variable AGE > Mean age, taking into account Fare, Embarked port, SibSP, Parch, Sex and Class, to imput missing NA

Rest of Missing values with mean according to Class only

```{r}
summary(Val$Age)

Val_2 <- merge(Val, mean_age_varios,by=c( "SibSp","Parch","sex_cod","Pclass"),all.x = TRUE)

Val_2$Age.x <- ifelse(is.na(Val_2$Age.x),Val_2$Age.y,Val_2$Age.x )

Val_2$Age.y <- NULL

colnames(Val_2 ) <- c("SibSp","Parch","sex_cod","Pclass" ,"PassengerId","Survived" ,"Name" , "Age" ,"Ticket", "Fare","Cabin" ,"Embarked" )

Val <- Val_2

#Se reordena el dataset
Val_2 <- cbind.data.frame(Val$PassengerId,Val$Survived,Val$Pclass,Val$Name,Val$sex_cod,Val$Age,Val$SibSp,Val$Parch,Val$Ticket,Val$Fare,Val$Cabin,Val$Embarked)

colnames(Val_2 ) <- c("PassengerId","Survived","Pclass"  ,"Name","sex_cod" , "Age" ,"SibSp","Parch","Ticket", "Fare","Cabin" ,"Embarked" )
Val <- Val_2

###REVIEW OF RESULTS

summary(Val$Age)

#Mean according to class
Val_2 <- merge(Val, mean_age_class,by='Pclass',all.x = TRUE)

Val_2$Age.x <- ifelse(is.na(Val_2$Age.x),Val_2$Age.y,Val_2$Age.x )

Val_2$Age.y <- NULL

Val <- Val_2

#Se reordena el dataset
Val_2 <- cbind.data.frame(Val$PassengerId,Val$Survived,Val$Pclass,Val$Name,Val$sex_cod,Val$Age,Val$SibSp,Val$Parch,Val$Ticket,Val$Fare,Val$Cabin,Val$Embarked)

colnames(Val_2 ) <- c("PassengerId","Survived","Pclass"  ,"Name","sex_cod" , "Age" ,"SibSp","Parch","Ticket", "Fare","Cabin" ,"Embarked" )
Val <- Val_2

#clean-up
rm(Val_2)

###REVIEW OF RESULTS

summary(Val$Age)
```

Analisis of varibale NAME, CABIN, TICKET

```{r}
Val$Name <-NULL
Val$Cabin <-NULL
Val$Ticket <- NULL
```

Analisis of variable Embarked
```{r}
table(Val$Embarked)
#Change to numeric
Val$Embarked <- as.numeric(Val$Embarked)

```



##Generation of different models to see which one fits better
Datasets to send to models
```{r}
Trn_m <- Trn
Trn_m$PassengerId <- NULL


Val_m <- Val
Val_m$PassengerId <- NULL
Val_m$Survived <-NULL
```

Lineal Regresion
```{r}
Rl <- glm(Survived ~., data=Trn_m)

rl_pred <- predict(Rl,newdata=Val_m,type = 'response')

rl_pred
#Matriz de confusion
table(Val$Survived, rl_pred > 0.5)

rl_pred <- ifelse(rl_pred > 0.5,1,0)

RL_results <- rmse(rl_pred, Val$Survived)
RL_results <- accuracy(Val$Survived,rl_pred)
Accuracy <-c('Reg Lineal','All', RL_results)

Results_3 <- c('Reg Lineal','All', RL_results)

#clean-up
rm(rl_pred)
```

Arbol de Decision
```{r}
A_T <- rpart(Survived~., data=Trn_m)

rpart.plot(A_T,fallen.leaves = F, type=2)

A_pred <- predict(A_T,Val_m)

table(Val$Survived, A_pred > 0.5)

Al_pred <- ifelse(A_pred > 0.5,1,0)

A_results <- rmse(A_pred, Val$Survived)
Results_2 <- c('Arbol','All', A_results)
Results <- rbind(Results,Results_2)

A_results <- accuracy(Val$Survived,Al_pred)
Results_3 <- c('Arbol','All', A_results)
Accuracy <- rbind(Accuracy,Results_2)

#clean-up
rm(A_pred)
```

Stepwise - Forward - Backward
```{r}
minimo <-glm(Survived ~ 1, data = Trn_m)

maximo <-glm(Survived ~ ., data = Trn_m)

#FORWARD
foward <- step(minimo, scope = list(lower=minimo, upper=maximo), direcction = 'foward')

F_pred <- predict(foward,Val_m)

table(Val$Survived, F_pred > 0.5)

F_pred <- ifelse(F_pred > 0.5,1,0)

F_results <- rmse(F_pred, Val$Survived)
Results_2 <- c('Forward','All', F_results)
Results <- rbind(Results,Results_2)

F_results <- accuracy(Val$Survived,F_pred)
Results_3 <- c('Forward','All', F_results)
Accuracy <- rbind(Accuracy,Results_3)


#BACKWARD
backward <- step(minimo, scope = list(lower=minimo, upper=maximo), direcction = 'backward')

B_pred <- predict(backward,Val_m)

table(Val$Survived, B_pred > 0.5)

B_pred <- ifelse(B_pred > 0.5,1,0)


B_results <- rmse(B_pred, Val$Survived)
Results_2 <- c('Backward','All', B_results)
Results <- rbind(Results,Results_2)

B_results <- accuracy(Val$Survived,F_pred)
Results_3 <- c('Backward','All', B_results)
Accuracy <- rbind(Accuracy,Results_3)

#STEPWISE
stepwise <- step(minimo, scope = list(lower=minimo, upper=maximo), direcction = 'stepwise')

S_pred <- predict(stepwise,Val_m)

table(Val$Survived, S_pred > 0.5)

S_pred <- ifelse(S_pred > 0.5,1,0)

S_results <- rmse(S_pred, Val$Survived)
Results_2 <- c('Stepwise','All', S_results)
Results <- rbind(Results,Results_2)

S_results <- accuracy(Val$Survived,F_pred)
Results_3 <- c('Stepwise','All', S_results)
Accuracy <- rbind(Accuracy,Results_3)

#clean-up
rm(B_pred)
rm(F_pred)
rm(S_pred)
```

Ridge y Lasso

```{r}
vExp <- c("Pclass","sex_cod" , "Age" ,"SibSp","Parch", "Fare","Embarked")

vObj <- c("Survived")

#Ridge
ridge <- funcRIDGE_LASSO(Trn_m,Val_m,0,vExp,vObj)
R_results <- rmse(ridge, Val$Survived)
Results_2 <- c('Ridge','All', R_results)
Results <- rbind(Results,Results_2)

R_results <- accuracy(Val$Survived,F_pred)
Results_3 <- c('Ridge','All', R_results)
Accuracy <- rbind(Accuracy,Results_3)

#Lasso
lasso <- funcRIDGE_LASSO(Trn_m,Val_m,1,vExp,vObj)
L_results <- rmse(lasso, Val$Survived)
Results_2 <- c('Lasso','All', L_results)
Results <- rbind(Results,Results_2)

L_results <- accuracy(Val$Survived,F_pred)
Results_3 <- c('Lasso','All', L_results)
Accuracy <- rbind(Accuracy,Results_3)

```

##Clean-up of TEST with learnt variables
Analisis of variable SEX > New numeric varible sex, 1 for male, 0 for female
```{r}
table(Test$Sex)

Test <- merge(Test, sex_cod,by='Sex')

Test$sex_cod <- unfactor(Test$sex_cod)

#Check of results
table(Test$Sex)
table(Test$sex_cod)

#Elimination of variable Sex
Test$Sex <-NULL

```

Analisis of variable AGE > Mean age, taking into account Fare, Embarked port, SibSP, Parch, Sex and Class, to imput missing NA

Rest of Missing values with mean according to Class only

```{r}
summary(Test$Age)

Test_2 <- merge(Test, mean_age_varios,by=c( "SibSp","Parch","sex_cod","Pclass"),all.x = TRUE)

Test_2$Age.x <- ifelse(is.na(Test_2$Age.x),Test_2$Age.y,Test_2$Age.x )

Test_2$Age.y <- NULL

colnames(Test_2 ) <- c("SibSp","Parch","sex_cod","Pclass" ,"PassengerId","Name" , "Age" ,"Ticket", "Fare","Cabin" ,"Embarked" )

Test <- Test_2

#Se reordena el dataset
Test_2 <- cbind.data.frame(Test$PassengerId,Test$Pclass,Test$Name,Test$sex_cod,Test$Age,Test$SibSp,Test$Parch,Test$Ticket,Test$Fare,Test$Cabin,Test$Embarked)

colnames(Test_2 ) <- c("PassengerId","Pclass"  ,"Name","sex_cod" , "Age" ,"SibSp","Parch","Ticket", "Fare","Cabin" ,"Embarked" )
Test <- Test_2

###REVIEW OF RESULTS

summary(Test$Age)

#Mean according to class
Test_2 <- merge(Test, mean_age_class,by='Pclass',all.x = TRUE)

Test_2$Age.x <- ifelse(is.na(Test_2$Age.x),Test_2$Age.y,Test_2$Age.x )

Test_2$Age.y <- NULL

Test <- Test_2

#Se reordena el dataset
Test_2 <- cbind.data.frame(Test$PassengerId,Test$Pclass,Test$Name,Test$sex_cod,Test$Age,Test$SibSp,Test$Parch,Test$Ticket,Test$Fare,Test$Cabin,Test$Embarked)

colnames(Test_2 ) <- c("PassengerId","Pclass"  ,"Name","sex_cod" , "Age" ,"SibSp","Parch","Ticket", "Fare","Cabin" ,"Embarked" )
Test <- Test_2

#clean-up
rm(Test_2)

###REVIEW OF RESULTS

summary(Test$Age)
```

Analisis of varibale NAME, CABIN, TICKET

```{r}
Test$Name <-NULL
Test$Cabin <-NULL
Test$Ticket <- NULL
```

Analisis of variable Embarked
```{r}
table(Test$Embarked)
#Change to numeric
Test$Embarked <- as.numeric(Test$Embarked)

```

##Generation of prediction for Test with best model
```{r}
#Foward
A_pred_final <- predict(foward,Test)

A_pred_final <- ifelse(A_pred_final > 0.5,1,0)

A_pred_final <- cbind(Test$PassengerId,A_pred_final)
summary(A_pred_final)

A_pred_final <- as.data.frame(A_pred_final)

generarArchivo(A_pred_final,"Prediction5_FowardAccuracy.csv")
```
#####################HASTA ACA##########


##Models with ONLY AGE, SEX AND CLASS
Datasets to send to models
```{r}
Trn_m <- Trn
Trn_m$PassengerId <- NULL
Trn_m$SibSp <- NULL
Trn_m$Parch <-NULL
Trn_m$Fare <-NULL
Trn_m$Embarked<-NULL
summary(Trn_m)

Val_m <- Val
Val_m$PassengerId <- NULL
Val_m$Survived <-NULL
Val_m$SibSp <- NULL
Val_m$Parch <-NULL
Val_m$Fare <-NULL
Val_m$Embarked<-NULL
summary(Val_m)

```
